%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{A Review: The Recovery of 3D Facial Shape From Images and Videos}\label{ch:background}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\minitoc{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The recovery of 3D facial shape from images and videos has a rich history
that is closely interwoven with the more general area of generic 3D shape
recovery. Due to the many applications of 3D facial data and the
relative ease of facial data collection, many generic methods for 3D shape
recovery have given examples of reconstructions for faces. Therefore, it
becomes difficult to separate methods that focus on 3D facial shape recovery
and more generic methods that provide facial shape recovery results
amongst many other examples. This is particularly prevalent in the Structure-
from-Motion (SfM) literature whereby few papers focus solely on face recovery
yet many papers show examples of facial reconstructions. In order to provide
a thorough treatment on both the history and the current state-of-the-art for
3D facial shape recovery, it is necessary to isolate key examples
of these more generic methodologies and present their application to faces.
Furthermore, there are examples of 3D facial shape recovery across all methods
of general shape recovery given in \cref{tbl:3d_recovery_methods} and thus
the relevant literature is vast. In this section, we will attempt to provide
an exhaustive review of methods for 3D facial shape recovery and will provide
literature for both sparse and dense shape recovery, despite the focus
of this thesis being on dense recovery. Given the scope of the review, the
literature will be separated into the categories provided by
\cref{tbl:3d_recovery_methods}. Each subsection will briefly describe the
key mechanics involved for the reconstruction method but will not focus on
an exhaustive review of all of the relevant literature for a given method.
Instead, only those methods that focus on facial shape recovery or provide
representative reconstructions of faces will be considered.

The rest of this section will proceed as follows. \cref{sec:bg_db_capture} will
discuss the currently available databases of 3D facial shape in the form
of meshes and reflectance information. It will also briefly touch on recent
innovations for capturing these databases. \cref{sec:bg_sfs} provides an overview
of Shape-from-Shading (SfS) methods for facial surface normal recovery from
single images. \cref{sec:bg_ps} discusses both calibrated and
Uncalibrated Photometric Stereo methods. When discussing Uncalibrated methods,
more recent large-scale methods from Internet images are also included.
\cref{sec:bg_3dmm} presents the relevant literature for Morphable Models,
beginning with work of \citet{volker1999morphable}.
\cref{sec:bg_model_based} gives an
overview of methods that attempt to directly infer shape information from images
by using models. \cref{sec:bg_sfm} discusses Structure-from-Motion methods,
including relevant methods that provide representative reconstructions
on faces despite not leveraging facial models.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{background/databases_and_capture.tex}
\input{background/shape_from_shading.tex}
\input{background/photometric_stereo.tex}
\input{background/morphable_models.tex}
\input{background/shape_transfer_model_based.tex}
\input{background/structure_from_motion.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\stopcontents[chapters]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%